Depression and suicidality as credible signals of need

### 2021-05-07 Paper.rmd
Finished modifying power stuff

```{r}
pwr80_90 <- 
  power_curve %>% 
  group_by(Base) %>% 
  summarise(
    pwr80 = min(sample_size[power>=.8]), 
    pwr90 = min(sample_size[power>=.9])
  )
```

We used the pilot data to estimate the sample sizes needed to detect an effect of the *Depression* signal vs. *Verbal request* control on *Belief* in the victim's need. Power was about 80% for a sample size of about `r pwr80_90[[2,2]]`, and was about 90% for a sample size of about `r pwr80_90[[2,3]]`. We also estimated the sample size needed to detect an effect of *Depression* vs. a *Control* condition that was meant to not include any signaling (but which we realized, post hoc, did include a verbal request). Power was about 80% for a sample size of about `r pwr80_90[[1,2]]`, and was about 90% for a sample size of about `r pwr80_90[[1,3]]`. See Figure \@ref(fig:power). Given our \$1500 USD budget, we aimed for a sample size of 124 for treatment plus control conditions, and 1240 for all conditions in the study. For more details, see the SI.

@@

**Thwarted marriage vignette**: Indian sample only. Each participant was asked to imagine that their family was trying to arrange a dowry for their older daughter (the signaler in this vignette) so she can marry a man she already loves, while still saving enough money for their younger daughter's dowry. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. Any increase in the dowry will come at the younger daughter's expense. The participant therefore proceeds to arrange a marriage to the second man as the first man's family makes arrangements with a different woman.

@@

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. Based on findings that a slider's starting position may bias results [@liu2019], we had concerns that participants would be less likely to move away from intermediate starting values than they would be in reality. For this reason, we set each T1 slider to start fully to the left (0). For the exact wording of the question and the labels on the sliders, see table 2.

@@

Consistent with the idea that depression and suicidality are credible signals of need, the two signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler relative to less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. Compared to these scenarios, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family members or a romantic partner).

@@

Overall, our results on the connection between belief in need and likelihood of help are consistent with credible signaling and bargaining approaches to depression and suicide. At T1, *Belief* and *Action* were highly correlated yet distinct; while at T2, the effect of signal on both *Belief* and *Action* was to increase the magnitude of pro-signaler responses, with the effect of signal on *T2Action* being largely mediated by signal's effect on belief in need. This relationship between belief in need and likelihood of support and the importance of signal cost should be expected in scenarios when a target of a signal of need would benefit from helping if the signaler's need was genuine but face substantial net costs if it was not. Importantly, such scenarios appear to be common among modern humans and have likely been since individuals could increase their inclusive fitness through supporting kin or important cooperative partners who could return to providing benefits to supporters once their situation returned to normal [@SugiyamaSugiyama2003; @tooby1996].

@@

Another limitation is that our control included a signal in the form of a verbal request, preventing comparisons to non-signaling strategies. Similarly, our design did not include vignettes with male signalers. For this reason, we have no data on the possibility of sex differences in the effectiveness of the signaling strategies examined. Although not predicted theoretically, such differences are possible if the costs of signaling vary between the sexes due to differential access to alternative bargaining strategies [@HagenRosenstrom2016] or if one sex suffered greater negative reputational effects due to emotions and behaviors in the vignettes. It is also possible the costliness of the situations presented in the vignettes differ by sex.

In terms of data quality our largest limitation is that, compared to the US sample, far more Indian participants were filtered out by our attention checks, something which could be consistent with botting, unfamiliarity with English, or low effort responses [@kennedy2020]. Assuming this corresponded with lower-quality responses among those who passed the attention checks, the weak signal effect in the thwarted marriage vignette may simply be due to added noise rather than differences in the scenario or the effectiveness of the signals compared to those in the US.

Slider anchor 
Regex: stopped in SI
T\d \w+
# Conclusion

@@

We estimated power using our pilot study data to compare the effect of the *Depression* signal vs. the *Verbal request* control on *Belief* (that the sister needed the money) at T2 controlling for *Belief* at T1 in a linear regression model. Specifically, we resampled the pilot data with replacement, fit the linear regression model, and extracted the p-value for the signal coefficient, repeating this 2000 times for each sample size ranging from N=20 to N=200. Power was computed as the proportion of p-values \< 0.05. Power was about 80% for a sample size of about `r pwr80_90[[2,2]]`, and was about 90% for a sample size of about `r pwr80_90[[2,3]]`. We then repeated this entire procedure to compare *Depression* to a (flawed) *Control* condition in the pilot study that was meant to not include any signaling (but which we realized, post hoc, did include a verbal request yet differed in substantive ways from the treatment conditions, and therefore was not suitable as a control). Power was about 80% for a sample size of about `r pwr80_90[[1,2]]`, and was about 90% for a sample size of about `r pwr80_90[[1,3]]`. See Figure \@ref(fig:power).

```{r power, fig.cap="Power curve for between-subjects effect of Depression vs. Verbal request signal on T2 Belief, controlling for T1 Belief, and Depression vs. a Control condition that was unique to the pilot study."}
  ggplot(power_curve, aes(sample_size, power, color=Base)) +
    geom_line() +
    labs(x = '\n Sample size', y = 'Power\n') +
    theme_minimal(15)
```

>>>>

```{r}
pwr80_90 <- 
  power_curve %>% 
  summarise(
    pwr80 = min(sample_size[power>=.8]), 
    pwr90 = min(sample_size[power>=.9])
  )
```

We used the pilot data to estimate the sample sizes needed to detect an effect of the *Mild depression* signal vs. *Verbal request* control on *Belief* in the victim's need. Power was about 80% for a sample size of about `r pwr80_90[[1, 1]]`, and was about 90% for a sample size of about `r pwr80_90[[1, 2]]`. See Figure \@ref(fig:power). Given our \$1500 USD budget, we aimed for a sample size of 120-130 for treatment plus control conditions, and 1200-1300 for all conditions in the study. For more details, see the SI.

@@

**Thwarted marriage vignette**: Indian sample only. Each participant was asked to imagine that their family was trying to arrange a dowry for their older daughter so she can marry a man she already loves, while still saving enough money for their younger daughter's dowry. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. Any increase in the dowry will come at the younger daughter's expense. The participant therefore proceeds to arrange a marriage to the second man as the first man's family makes arrangements with a different woman.

@@

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. Based on findings that a slider's starting position may bias results [@liu2019], we had concerns that participants would be less likely to move away from intermediate starting values than they would be in reality. For this reason, we set each T1 slider to start fully to the left (0). For the exact wording of the question and the labels on the sliders, see table 1.

@@

Consistent with costly signaling theory, costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. Compared to these scenarios, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family members or a romantic partner).

@@

Overall, our results on the connection between belief in need and likelihood of help are consistent with costly signaling and bargaining approaches to depression and suicide. At T1, *Belief* and *Action* were highly correlated yet distinct; while at T2, the effect of signal on both *Belief* and *Action* was to increase the magnitude of pro-signaler responses, with the effect of signal on *T2Action* being largely mediated by signal's effect on belief in need. This relationship between belief in need and likelihood of support and the importance of signal cost should be expected in scenarios when a target of a signal of need would benefit from helping if the signaler's need was genuine but face substantial net costs if it was not. Importantly, such scenarios appear to be common among modern humans and have likely been since individuals could increase their inclusive fitness through supporting kin or important cooperative partners who could return to providing benefits to supporters once their situation returned to normal [@SugiyamaSugiyama2003; @tooby1996].

@@

<!-- However, the fact that the relationships among signal cost, belief in need, and likelihood of providing support all largely fit existing theory (narrow and broad) indicates that participants paid attention to the relevant details and responded in ways that suggest some degree of external validity. For this reason, the benefits of addressing causal relationships that came through our experimental vignette approach likely outweighed the costs. -->

Another limitation is that, compared to the US sample, far more Indian participants were filtered out by our attention checks, something which could be consistent with botting, unfamiliarity with English, or low effort responses [@kennedy2020]. Assuming this corresponded with lower-quality responses among those who passed the attention checks, the weak signal effect in the thwarted marriage vignette may simply be due to added noise rather than differences in the scenario or the effectiveness of the signals compared to those in the US.

# Conclusion

@@

We estimated power using our pilot study data to compare the effect of the *Mild depression* signal vs. the *Verbal request* control on *Belief* (that the sister needed the money) at T2 controlling for *Belief* at T1 in a linear regression model. Specifically, we resampled the pilot data with replacement, fit the linear regression model, and extracted the p-value for the signal coefficient, repeating this 2000 times for each sample size ranging from N=20 to N=200. Power was computed as the proportion of p-values \< 0.05. Power was about 80% for a sample size of about `r pwr80_90[[1,1]]`, and was about 90% for a sample size of about `r pwr80_90[[1,2]]`. See Figure \@ref(fig:power).

```{r power, fig.cap="Power curve for between-subjects effect of Mild depression vs. Verbal request signal on T2 Belief, controlling for T1 Belief."}
  ggplot(power_curve, aes(sample_size, power)) +
    geom_line() +
    labs(x = '\n Sample size', y = 'Power\n') +
    theme_minimal(15) + theme(axis.title.y = element_text(angle = 0))
```
